---
title: "The Benefits of Immigration in the USA"
subtitle: ""
author: "Ghislain Nono Gueye, Ph.D."
date: "10/30/2018"
output:
  xaringan::moon_reader:
    css: [default, "style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
```

class: inverse, middle, center

# Analyzing welfare

---

### Costs and benefits

- Every economic choice has benefits and costs.

- Rationality: making choices when benefits outweigh costs.

<br/>

- [E.g.1] The guns and butter trade-off.

- [E.g.2] Industrialization vs clean environment

- [E.g.3] Efficiency vs Equality

---

class: inverse, middle, center

# Immigration surplus

---

### Factors of production

- .red[Capital]
    
- Entrepreneurship
    
- Land
    
- .red[Labor]

<br/>

--

.center[
Capital + Labor = Production $\rightarrow$ Income
]

<br/>

--

- .blue[Capital in the USA is under-utilized in many industries.]

--

- .red[Immigration surplus] is the increase in the aggregate wealth of the natives resulting from the productive contributions of immigrants.

--

- .red[Estimate: 0.2 to 0.4% of GDP ($36 to $72 billion per year)]

---

class: inverse, middle, center

# Relieving bottlenecks/shortages

---

**Immigrants naturally flow to where workers are needed.**

--

.blue[Example 1]

During World War II, Mexican immigrants helped alleviate shortages arising from the war effort.

--

.blue[Example 2]

During the oil boom of late 1970s and early 1980s, there was a record flow of immigrants to Texas.

--

.blue[Example 3]

The high-tech industry received many immigrants during the internet boom  in the 2000s.

---

class: inverse, middle, center

# Increase in overall efficiency/productivity

---

### Some statistics

**The average immigrant in the USA innovates more than the average native.**

- .blue[Immigrants are concentrated in STEM fields:]

    - 44% of medical scientists
    
    - 42% of computer software developers
    
    - 35% of faculty in engineering
    
    - 39% of faculty in computer science
    
--

- .blue[The US has more immigrant inventors than every other country combined:]

---

```{r immigrant_inventors1, message = FALSE, warning = FALSE, echo = FALSE, out.width = "75%", fig.align = "center"}

dat <- read_csv("data/immigrant_inventors_by_country.csv") %>%
	rename(patents = `Immigrants with patents`) %>%
	mutate(country = fct_reorder(country, patents),
			 patents = patents/1000)

ggplot(data = dat, aes(x = country, y = patents, color = country)) +
	geom_point() +
	geom_linerange(aes(ymin = 0, ymax = patents)) +
	labs(x = "", y = "", title = "Number of patents owned by immigrants between 2000 and 2016", subtitle = "(in thousands)") +
	theme_bw() +
	theme(legend.position = "none") +
	coord_flip()
```

---

```{r immigrant_inventors2, echo = FALSE, out.width = "85%", fig.align = "center"}
dat %>%
	mutate(country2 = if_else(country == "United States", "United States", "All other countries"),
			 country2 = as.factor(country2)) %>%
	group_by(country2) %>%
	summarize(patents = sum(patents)) %>%
	mutate(country2 = fct_reorder(country2, desc(patents))) %>%
	
	ggplot(aes(x = country2, y = patents, fill = country2)) +
	geom_bar(stat = "identity", show.legend = FALSE) +
	labs(x = "", y = "", title = "Number of patents owned by immigrants between 2000 and 2016", subtitle = "(in thousands)") +
	theme_bw()
```

---

class: inverse, middle, center

# THANK YOU!


